---
title: "respR_analysis"
author: "Mike Connelly"
date: "6/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/mikeconnelly/computing/projects/anti_phys")
```
```{r packages}
library("tidyverse")
library("purrr")
library("respR") # devtools::install_github("januarharianto/respR")
library("rstatix")
library("modelr")
```
```{r colors}
#rename to named colors
treatcolors <- c("lightblue", "dodgerblue", "seagreen1")
#treatfills <- c()
#shapes <- c(18, 9, 16, 10) # may be useful for genotype variables: Gulf, Location, mtORF, symbiont, etc.
genotype_colors <- read_csv("data/genotype_colors.csv") # color names csv corresponding to excel sheet colors
# default ggplot fill may also work in this case
genocolors <- c(genotype_colors$color, "grey")
```
```{r theme}
theme_set(theme_bw())
```
```{r fragments}
fragments <- read_csv("data/fragments.csv", col_types = cols(
  Fragment_ID = col_character()
))
# factor levels
treatment_levels <- c("Baseline", "Control", "Antibiotics")
sampling_levels <- c("Day 0", "Day 7")
# sampling_levels <- 
genotypes <- read_csv("data/genotypes.csv")
genotype_levels <- c("PAN-5", "PAN-10", "PAN-32", "PAN-34", "PAN-35", "PAN-37", "PAN-39", "PAN-41", "PAN-43","PAN-44", "PAN-78", "PAN-79", "PAN-83", "URA-51", NA)
gulf_levels <- c("Chiriqui", "Panama")
loc_levels <- c("Uva", "Saboga", "Uraba")
mtorf_levels <- c("Type_1", "Type_2", "Type_3", "unknown")
sym_levels <- c("C1bc", "C1d", "D1", "D2", "unknown")
```
```{r}
# read in fragment mass data
mass <- read_csv("data/fragment_mass.csv", col_types = cols(
  Fragment_ID = col_character(),
  Mass = col_double()
))
```
```{r fragment_mass_analysis}
fragment_mass <- left_join(fragments, mass, by = "Fragment_ID")
fragment_mass$Treatment <- factor(fragments$Treatment, levels = treatment_levels, ordered = TRUE)
fragment_mass$Genotype <- factor(fragments$Genotype, levels = genotype_levels, ordered = TRUE)
# 
fragment_mass_boxplot <- fragment_mass %>% 
  filter(Treatment != "Baseline") %>% 
  ggplot(aes(x = Genotype, y = Mass, fill = Genotype)) +
  geom_boxplot() +
  geom_jitter(shape = 21) +
  # facet_grid(~Treatment, scales = "free") +
  scale_fill_manual(values = genocolors)
fragment_mass_boxplot
# ggsave("outputs/figures/fragment_mass_experiment_boxplot.pdf", fragment_mass_boxplot,  device = "pdf", height = 4, width = 10)
# print(fragment_mass_boxplot)
```

## Import plate fragment-position keys
```{r plate_keys}
# Fragment positions in the 24-well plates were randominzed each day during media changes, and positions were maintained through the IPAM assessment, microscopy imaging, and microplate respirometry
# Information about which fragments or blank wells were in which position needs to be imported and made to correspond with the appropriate AOI information (IPAM) and well position information (respirometry) in the imported source data

# import IPAM area of interest (AOI) to fragment ID keys to later connect measurements to sample metadata
plate_keys <- list.files(path = "./data/plate_keys", pattern = "plate_key_d[[:digit:]]\\.csv", full.names = T)
# check to ensure all files are present
plate_keys

# Create lookup table for AOI_ID from Well_ID
wells_aoi <- c("A1" = 1, "A2" = 2, "A3" = 3, "A4" = 4, "A5" = 5, "A6" = 6,
               "B1" = 7, "B2" = 8, "B3" = 9, "B4" = 10, "B5" = 11, "B6" = 12,
               "C1" = 13, "C2" = 14, "C3" = 15, "C4" = 16, "C5" = 17, "C6" = 18,
               "D1" = 19, "D2" = 20, "D3" = 21, "D4" = 22, "D5" = 23, "D6" = 24)
aoi_levels <- as.character(1:24)
# Import all plate keys 
plate_key_master <- plate_keys %>%
  map_dfr(read_csv) %>% 
  select(Meas_Day, Sample_ID, Fragment_ID, Plate_Well_ID) %>% 
  arrange(Sample_ID) %>% 
  separate(col = Plate_Well_ID, into = c("Plate", "Well"), sep = "-")

# add column that converts the well ID information into AOI for IPAM
plate_key_master$AOI <- as.character(unname(wells_aoi[plate_key_master$Well]))
# rbind into one dataframe?

plate_key_master <- plate_key_master %>% 
  unite(AOI_ID, Plate, AOI, sep = "_", remove = FALSE) %>% 
  unite(AOI_ID_key, Meas_Day, AOI_ID, sep = "_", remove = FALSE) %>% 
  unite(Well_ID, Plate, Well, sep = "_", remove = FALSE) %>% 
  unite(Well_ID_key, Meas_Day, Well_ID, sep = "_", remove = FALSE) %>% 
  arrange(Well_ID_key)
```

I want to devise a way to import sample and well metadata to help analyse the rates... done!
I will need to include sample polyp count, weight and/or volume displacement to standardize the rates
It may be necessary to make even *smaller* fragments, like just the branch tips.
I need to record the units that the respirometer is measuring in... or change them to match respR.

## Import respirometry data files
```{r}
# ?import_file()
# List all plate information files in the target import directory
resp_data_files <- list.files(path = "./data/resp", pattern = "*.xlsx", full.names = T)
# name files for easier parsing of path contents including measurement day (day1 - day7), plate number (plate1 - plate6), and date (MMDDYY)
# example path: "./data/resp/day1_plate1_083120.xlsx
resp_data_files
```
```{r}
# need to figure out how to iteratively import all the plate files for a given day
# day2_plate6 <- import_file(path = resp_data_files[6])
# 
plate_names <- gsub("./data/resp/", "", resp_data_files)
plate_names_1 <- gsub("_[0-9]{6}_Oxygen.xlsx", "", plate_names)
resp_all <- map(resp_data_files, import_file) %>% set_names(plate_names_1)
resp_all_df <- map_df(resp_data_files, import_file)
# 
resp_day2 <- map(resp_data_files[1:6], import_file)
resp_day3 <- map(resp_data_files[7:12], import_file)
resp_day4 <- map(resp_data_files[13:18], import_file)
resp_day5 <- map(resp_data_files[19:24], import_file)
```

```{r}
# unit_args()
# O2 unit from the microplate respirometer instrument: "%"
# Time unit: "s"/"m"
# Output mass unit: "ug"
```
```{r}
names(resp_all)
resp_all$day2_plate1
```

## 24-channel plate plots
```{r single_plate_24_channel}
plate <- resp_day4[[2]]
# 
channel_plot <- plate %>% 
  filter(`Time/Min.` > 10) %>% # filter out time before 10 minutes, this is usually how long the plate takes to equilibrate
  select(`Time/Min.`, A1:D6) %>% 
  pivot_longer(cols = A1:D6,
               names_to = "Well",
               values_to = "pct_oxygen") %>% 
  ggplot(aes(x = `Time/Min.`, y = pct_oxygen)) +
  geom_point(aes(color = Well))
channel_plot
```
```{r daily_plates_24_channel}
pdf("./outputs/figures/day5_plates_24-channel.pdf")
for (i in seq_along(resp_day5)) {
plate <- resp_day5[[i]]

plot_title <- paste("Respirometer oxygen consumption plate", i, "24-channels")

channel_plot <- plate %>% 
  filter(`Time/Min.` > 10) %>% # filter out time before 10 minutes, this is usually how long the plate takes to equilibrate
  select(`Time/Min.`, A1:D6) %>% 
  pivot_longer(cols = A1:D6,
               names_to = "Well",
               values_to = "pct_oxygen") %>% 
  ggplot(aes(x = `Time/Min.`, y = pct_oxygen)) +
  geom_point(aes(color = Well)) +
  ggtitle(plot_title)
print(channel_plot)
}
dev.off()
```
```{r}
resp_all_df <- bind_rows(resp_all, .id = "id") %>% separate(id, into = c("day", "plate"), sep = "_")
```
```{r}
raw_data <- resp_all_df %>% 
  filter(`Time/Min.` > 10) %>%
  filter(`Time/Min.` < 60) %>% # filter out time before 10 minutes, this is usually how long the plate takes to equilibrate
  select(day, plate, `Time/Min.`, A1:D6) %>% 
  pivot_longer(cols = A1:D6,
               names_to = "Well",
               values_to = "pct_oxygen") %>% 
  ggplot(aes(x = `Time/Min.`, y = pct_oxygen)) +
  geom_point(aes(color = Well)) +
  facet_grid(plate ~ day, switch = "y")
ggsave("./outputs/figures/raw_data_plates.pdf", raw_data, height = 6, width = 10)
```


## respR rate analysis
```{r respR single}
plate <- resp_day2[[2]]
# 

# Create empty data frame to hold output from rate calculation loop
rates_frame <- data.frame()
# Empty list structure to hold all rate objects
rates_list <- list()

plate %>%
  inspect(time = 2, oxygen = 4) 

# auto_rate
rate_test <- day2_plate6 %>% 
  filter(`Time/Min.` > 15) %>% # filter out time before 10 minutes, this is usually how long the plate takes to equilibrate
  inspect(time = 2, oxygen = 4) %>% 
  auto_rate() %>% 
  print()  %>%   # just a quick preview
  convert_rate(o2.unit = "%",
               time.unit = "min", 
               output.unit = "ug/m", 
               volume = 0.6, S = 35, t = 26, P = 0.977)
rate_test$output[1]

names(wells_aoi[1])
names(wells_aoi[4])
```
```{r plate}
# The below code works now! DO NOT CHANGE
# loop through rate calculation steps 
rates_frame <- data.frame("Well" = vector(), "rate" = numeric())
for (i in 3:26) {
rate <- plate %>%
  filter(`Time/Min.` > 15 & `Time/Min.` < 45) %>% # filter out time before 10 minutes, this is usually how long the plate takes to equilibrate
  # select() %>%
  inspect(time = 2, oxygen = i) %>% 
  # Incorporate background rate detection into the analysis
  # calc_rate() %>% 
  auto_rate() %>% # automatically determine most linear segment
  print()  %>%   # just a quick preview
  convert_rate(o2.unit = "%",
               time.unit = "min", 
               output.unit = "ug/m", 
               volume = 0.6, S = 35, t = 26, P = 0.977)
rates_frame[i-2, 1] <- names(wells_aoi[i-2])
rates_frame[i-2, 2] <- rate$output[1]
rates_list[i-2] <- rate
}
```

```{r respR_functions}
# Create a function that calculates the fragment oxygen consumption rate for each well of each plate on any given day
# Different versions for calc_rate() versus auto_rate() commands
auto_plate_resp <- function(plate) {
# Create empty data frame to hold output from rate calculation loop
rates_frame <- data.frame("Well" = vector(), "rate" = numeric())
# loop through rate calculation steps 
for (i in 3:26) {
rate <- plate %>%
  filter(`Time/Min.` > 10) %>% # filter out time before 10 minutes, this is usually how long the plate takes to equilibrate
  # select(2,i) %>%
  inspect(time = 2, oxygen = i) %>% 
  # Incorporate background rate detection into the analysis
  auto_rate() %>% # automatically determine most linear segment
  print()  %>%   # just a quick preview
  convert_rate(o2.unit = "%",
               time.unit = "min", 
               output.unit = "ug/m", 
               volume = 0.6, S = 35, t = 26, P = 0.977)
rates_frame[i-2, 1] <- names(wells_aoi[i-2])
rates_frame[i-2, 2] <- rate$output[1]
# colnames(rates_frame) <- c("Well", "rate")
}
return(rates_frame)
}
```
```{r}
rates_frame_auto <- auto_plate_resp(plate)
rates_frame_auto_2 <- rbind(plate_resp(plate), rates_frame_auto)
```


```{r}
# rates_frame_full <- data.frame()
# for (i in seq_along(resp_day3)) {
# plate <- resp_day3[[i]]
# plate_resp(plate)
# rates_frame_full <- rbind(rates_frame, rates_frame_full)
# }
# 
# Create empty data frame to hold output from rate calculation loop
rates_frame_mapped <- data.frame("Well" = vector(), "rate" = numeric())
rates_frame_auto_mapped <- resp_all %>% map_dfr(auto_plate_resp)
# attach metadata from keys
day_plate <- plate_key_master %>% 
  filter(Meas_Day %in% c("day2", "day3", "day4", "day5")) %>% 
  arrange(Well_ID_key)
# 
rates_frame_auto_key <- cbind(rates_frame_auto_mapped, day_plate[,1:8])
rates_frame_auto_meta <- rates_frame_auto_key %>%
  left_join(fragments, by = "Fragment_ID") %>%
  left_join(mass, by = "Fragment_ID") %>% 
  left_join(genotypes, by = "Genotype") %>% 
  mutate(rate_norm = -1*(rate / Mass))

rates_frame_auto_meta$Treatment <- factor(rates_frame_auto_meta$Treatment, levels = treatment_levels, ordered = TRUE)
rates_frame_auto_meta$Genotype <- factor(rates_frame_auto_meta$Genotype, levels = genotype_levels, ordered = TRUE)
```

## Exploratory analysis
```{r daily_rates_qc}
rates_frame_auto_meta %>% 
  ggplot(aes(Well, (rate)*-1)) +
  geom_boxplot()
rates_frame_auto_meta %>% 
  ggplot(aes(Well, (rate)*-1)) +
  geom_boxplot() +
  facet_grid(Meas_Day~.)
rates_frame_auto_meta %>% 
  ggplot(aes(Plate, (rate)*-1)) +
  geom_boxplot()
rates_frame_auto_meta %>% 
  ggplot(aes(Plate, (rate)*-1)) +
  geom_boxplot() +
  facet_grid(Meas_Day~.)
```
```{r daily_raw_rates_genotype}
daily_rates_boxplot <- rates_frame_auto_meta %>% 
  filter(Meas_Day == "day5") %>% 
  ggplot(aes(Genotype, (rate)*-1, fill = Genotype)) +
  geom_boxplot(aes(color = Treatment)) +
  geom_point(shape = 21) +
  # ylim(c(0.0, 0.8)) +
  scale_fill_manual(values = genocolors) + 
  scale_color_manual(values = treatcolors[2:3]) +
  ylab("raw oxygen consumption rate (ug/min)")
daily_rates_boxplot
ggsave("./outputs/figures/daily_rates_boxplot_day5.pdf", daily_rates_boxplot, device = "pdf", height = 5, width = 10)
```
```{r}
rates_frame_auto_meta %>% 
  ggplot(aes(Treatment, (rate)*-1)) +
  geom_boxplot(aes(color = Treatment)) +
  geom_point(shape = 21) +
  # ylim(c(0.0, 0.8)) +
  facet_grid(~ITS2) +
  scale_fill_manual(values = genocolors) + 
  scale_color_manual(values = treatcolors[2:3])
```
How do I normalize the rates for each individual microfragment?
How will these rates be affected by genotypic differences in morphology, density, etc.?
```{r daily_normalized_rates_genotype}
daily_norm_rates_boxplot <- rates_frame_auto_meta %>% 
  filter(Meas_Day == "day5") %>% 
  filter(Genotype != "NA") %>% 
  ggplot(aes(Genotype, rate_norm, fill = Genotype)) +
  geom_boxplot(aes(color = Treatment)) +
  geom_point(shape = 21) +
  # ylim(c(0.0, 0.8)) +
  scale_fill_manual(values = genocolors) + 
  scale_color_manual(values = treatcolors[2:3]) +
  ylab("mass-normalized oxygen consumption rate (ug/min/g)")
daily_norm_rates_boxplot
ggsave("./outputs/figures/daily_norm_rates_boxplot_day5.pdf", daily_norm_rates_boxplot, device = "pdf", height = 5, width = 10)
```

```{r progression_normalized_rates}

rates_frame_auto_meta %>% 
  filter(Genotype != "NA") %>% 
  ggplot(aes(Genotype, rate_norm, fill = Genotype)) +
  geom_boxplot(aes(color = Treatment)) +
  geom_point(shape = 21) +
  coord_cartesian(ylim = c(0, 400)) +
  facet_grid(Meas_Day~.) +
  scale_fill_manual(values = genocolors) + 
  scale_color_manual(values = treatcolors[2:3]) +
  ylab("mass-normalized oxygen consumption rate (ug/min/g)")
```
```{r}
rates_frame_auto_meta %>% 
  filter(!is.na(Fragment_ID)) %>%
  # filter(Gulf == "Panama") %>%
  ggplot(aes(Meas_Day, rate_norm)) +
  geom_boxplot(aes(fill = Genotype, color = Treatment)) +
  geom_point(aes(color = Treatment)) +
  geom_path(aes(group = Fragment_ID)) +
  # facet_wrap(~Genotype + Treatment) +
  facet_grid(Treatment~.)+
  scale_fill_manual(values = genocolors) + 
  scale_color_manual(values = treatcolors[2:3])
```

```{r}
gulf_progression_boxplot <- rates_frame_auto_meta %>% 
  filter(!is.na(Fragment_ID)) %>%
  # filter(Gulf == "Chiriqui") %>%
  # filter(Gulf == "Panama") %>%
  ggplot(aes(Meas_Day, rate_norm)) +
  geom_boxplot(aes(fill = Genotype, color = Treatment)) +
  geom_point(aes(color = Treatment)) +
  geom_path(aes(group = Fragment_ID)) +
  facet_wrap(~Genotype + Treatment) +
  # facet_grid(Treatment~.)+
  scale_fill_manual(values = genocolors) + 
  scale_color_manual(values = treatcolors[2:3])
gulf_progression_boxplot
ggsave("./outputs/figures/gulf_respiration_boxplot.pdf", height = 8, width = 10, device = "pdf")
```

```{r}
rates_geno <- rates_frame_auto_meta %>%
  group_by(Treatment, Genotype, Meas_Day) %>% 
  summarise(geno_rate_norm = mean(rate_norm)) %>% 
  filter(!is.na(Genotype))

geno_rates_plot <- rates_frame_auto_meta %>% 
  filter(Genotype != "NA") %>% 
  ggplot(aes(Meas_Day, rate_norm)) +
  geom_boxplot(aes(fill = Treatment)) +
  geom_path(data = rates_geno, aes(Meas_Day, geno_rate_norm, group = Genotype, color = Genotype)) +
  # geom_smooth(aes(Meas_Day, rate_norm, group = Genotype, color = Genotype)) +
  geom_point(data = rates_geno, aes(Meas_Day, geno_rate_norm, color = Genotype), size = 5) +
  coord_cartesian(ylim = c(0, 400)) +
  facet_grid(~Treatment) +
  scale_color_manual(values = genocolors) + 
  scale_fill_manual(values = treatcolors[2:3]) +
  ylab("mass-normalized oxygen consumption rate (ug/min/g)") + 
  ggtitle("Mass-normalized oxygen consumption rates during experiment")
geno_rates_plot
ggsave("./outputs/figures/geno_rates_boxplot.pdf", height = 5, width = 10, device = "pdf")
```

## Statistical analysis
```{r}
rates_frame_auto_meta1 <- rates_frame_auto_meta %>%
  filter(Sample_ID != "b") %>% 
  mutate(lrate = log10(rate_norm),
         lnrate = log(rate_norm, base = exp(1)),
         invrate = 1/rate_norm)

rates_frame_auto_meta1$Genotype
rates_frame_auto_meta1$Meas_Day <- factor(rates_frame_auto_meta1$Meas_Day, levels = c("day2", "day3", "day4", "day5"), ordered = TRUE)

shapiro.test(rates_frame_auto_meta1$rate)
ggdensity(rates_frame_auto_meta1$rate)
ggqqplot(rates_frame_auto_meta1$rate)

shapiro.test(rates_frame_auto_meta1$rate_norm)
ggdensity(rates_frame_auto_meta1$rate_norm)
ggqqplot(rates_frame_auto_meta1$rate_norm)

shapiro.test(rates_frame_auto_meta1$lnrate)
ggdensity(rates_frame_auto_meta1$lnrate) + stat_overlay_normal_density(color = "red", linetype = "dashed")
ggqqplot(rates_frame_auto_meta1$lrate)

shapiro.test(rates_frame_auto_meta1$invrate)
ggdensity(rates_frame_auto_meta1$invrate) + stat_overlay_normal_density(color = "red", linetype = "dashed")
ggqqplot(rates_frame_auto_meta1$invrate)
```
```{r}
rates_frame_auto_meta1 %>% 
  group_by(Genotype) %>% 
  shapiro_test(rate_norm, lnrate) %>% filter(p < 0.05) %>% View()
# %>% group_by(variable) %>% count()
```

```{r}
shapiro.test(mass$Mass)
ggdensity(mass$Mass)
```
```{r}
kruskal.test(rates_frame_auto_meta1)
```

```{r}
fit <- lm(rate_norm ~ Treatment + Genotype + Meas_Day, data = rates_frame_auto_meta1)
summary(fit)
coef(resp_mod)
# 
# Other useful functions
coefficients(fit) # model coefficients
confint(fit, level=0.95) # CIs for model parameters
fitted(fit) # predicted values
residuals(fit) # residuals
anova(fit) # anova table
vcov(fit) # covariance matrix for model parameters
influence(fit) # regression diagnostics
# diagnostic plots
layout(matrix(c(1,2,3,4),2,2)) # optional 4 graphs/page
plot(fit)
# 
shapiro.test(resid(fit))
```
```{r}
# Stepwise Regression
library(MASS)
fit <- lm(lrate ~ Treatment + Gulf + Genotype + Meas_Day, data = rates_frame_auto_meta1)
step <- stepAIC(fit, direction="both")
step$anova # display results
```

```{r}
# Global test of model assumptions
library(gvlma)
gvmodel <- gvlma(fit)
summary(gvmodel)
```

```{r}
library(modelr)
model_matrix(rates_frame_auto_meta1, rate_norm ~ Treatment + Gulf + Genotype + Meas_Day) %>% View()
```

