---
title: "respR_analysis"
author: "Mike Connelly"
date: "6/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/mikeconnelly/computing/projects/anti_phys")
```
```{r packages}
library("tidyverse")
library("purrr")
library("respR") # devtools::install_github("januarharianto/respR")
# library("rstatix")
# library("modelr")
# library("MASS")
# library("gvlma")
```
```{r colors}
#rename to named colors
treatcolors <- c("lightblue", "dodgerblue", "seagreen1")
#treatfills <- c()
#shapes <- c(18, 9, 16, 10) # may be useful for genotype variables: Gulf, Location, mtORF, symbiont, etc.
genotype_colors <- read_csv("data/genotype_colors.csv") # color names csv corresponding to excel sheet colors
# default ggplot fill may also work in this case
genocolors <- c(genotype_colors$color, "grey")
```
```{r theme}
theme_set(theme_bw())
```
```{r fragments}
fragments <- read_csv("data/fragments.csv", col_types = cols(
  Fragment_ID = col_character()
))
# factor levels
treatment_levels <- c("Baseline", "Control", "Antibiotics")
sampling_levels <- c("Day 0", "Day 7")
day_levels <- c("day2", "day3", "day4", "day5")
genotypes <- read_csv("data/genotypes.csv")
genotype_levels <- c("PAN-5", "PAN-10", "PAN-32", "PAN-34", "PAN-35", "PAN-37", "PAN-39", "PAN-41", "PAN-43","PAN-44", "PAN-78", "PAN-79", "PAN-83", "URA-51", NA)
gulf_levels <- c("Chiriqui", "Panama")
loc_levels <- c("Uva", "Saboga", "Uraba")
mtorf_levels <- c("Type_1", "Type_2", "Type_3", "unknown")
sym_levels <- c("C1bc", "C1d", "D1", "D2", "unknown")
```
```{r fragment_mass}
# read in fragment mass data
mass <- read_csv("data/fragment_mass.csv", col_types = cols(
  Fragment_ID = col_character(),
  Mass = col_double()
))
```
```{r fragment_mass_boxplot}
fragment_mass <- left_join(fragments, mass, by = "Fragment_ID")
fragment_mass$Treatment <- factor(fragments$Treatment, levels = treatment_levels, ordered = TRUE)
fragment_mass$Genotype <- factor(fragments$Genotype, levels = genotype_levels, ordered = TRUE)
# 
fragment_mass_boxplot <- fragment_mass %>% 
  filter(Treatment != "Baseline") %>% 
  ggplot(aes(x = Genotype, y = Mass, fill = Genotype)) +
  geom_boxplot() +
  geom_jitter(shape = 21) +
  # facet_grid(~Treatment, scales = "free") +
  scale_fill_manual(values = genocolors)
fragment_mass_boxplot
# ggsave("outputs/figures/fragment_mass_experiment_boxplot.pdf", fragment_mass_boxplot,  device = "pdf", height = 4, width = 10)
# print(fragment_mass_boxplot)
```

## Import plate fragment-position keys
```{r plate_keys, message = FALSE}
# Fragment positions in the 24-well plates were randominzed each day during media changes, and positions were maintained through the IPAM assessment, microscopy imaging, and microplate respirometry
# Information about which fragments or blank wells were in which position needs to be imported and made to correspond with the appropriate AOI information (IPAM) and well position information (respirometry) in the imported source data

# import IPAM area of interest (AOI) to fragment ID keys to later connect measurements to sample metadata
plate_keys <- list.files(path = "./data/plate_keys", pattern = "plate_key_d[[:digit:]]\\.csv", full.names = T)
# check to ensure all files are present
plate_keys

# Create lookup table for AOI_ID from Well_ID
wells_aoi <- c("A1" = 1, "A2" = 2, "A3" = 3, "A4" = 4, "A5" = 5, "A6" = 6,
               "B1" = 7, "B2" = 8, "B3" = 9, "B4" = 10, "B5" = 11, "B6" = 12,
               "C1" = 13, "C2" = 14, "C3" = 15, "C4" = 16, "C5" = 17, "C6" = 18,
               "D1" = 19, "D2" = 20, "D3" = 21, "D4" = 22, "D5" = 23, "D6" = 24)
aoi_levels <- as.character(1:24)
# Import all plate keys 
plate_key_master <- plate_keys %>%
  map_dfr(read_csv) %>% 
  dplyr::select(Meas_Day, Sample_ID, Fragment_ID, Plate_Well_ID) %>% 
  arrange(Sample_ID) %>% 
  separate(col = Plate_Well_ID, into = c("Plate", "Well"), sep = "-")

# add column that converts the well ID information into AOI for IPAM
plate_key_master$AOI <- as.character(unname(wells_aoi[plate_key_master$Well]))

# rbind into one dataframe
plate_key_master <- plate_key_master %>% 
  unite(AOI_ID, Plate, AOI, sep = "_", remove = FALSE) %>% 
  unite(AOI_ID_key, Meas_Day, AOI_ID, sep = "_", remove = FALSE) %>% 
  unite(Well_ID, Plate, Well, sep = "_", remove = FALSE) %>% 
  unite(Well_ID_key, Meas_Day, Well_ID, sep = "_", remove = FALSE) %>% 
  arrange(Well_ID_key)
```

I want to devise a way to import sample and well metadata to help analyse the rates... done!
I will need to include sample polyp count, weight and/or volume displacement to standardize the rates
It may be necessary to make even *smaller* fragments, like just the branch tips.
I need to record the units that the respirometer is measuring in... or change them to match respR.

## Import respirometry data files
```{r resp_files}
# ?import_file()
# List all plate information files in the target import directory
resp_data_files <- list.files(path = "./data/resp", pattern = "*.xlsx", full.names = T)
# name files for easier parsing of path contents including measurement day (day1 - day7), plate number (plate1 - plate6), and date (MMDDYY)
# example path: "./data/resp/day1_plate1_083120.xlsx
resp_data_files
```
```{r plate_names}
plate_names <- gsub("./data/resp/", "", resp_data_files)
plate_names <- gsub("_[0-9]{6}_Oxygen.xlsx", "", plate_names)
```
```{r respR_import_daily}
resp_day2 <- map(resp_data_files[1:6], import_file)
resp_day3 <- map(resp_data_files[7:12], import_file)
resp_day4 <- map(resp_data_files[13:18], import_file)
resp_day5 <- map(resp_data_files[19:24], import_file)
```
```{r respR_import_all}
resp_all <- map(resp_data_files, import_file) %>% set_names(plate_names)
# check to see if data file names are included for each plate
names(resp_all)
# resp_all_df <- map_df(resp_data_files, import_file)
# 
```
```{r}
unit_args()
# O2 unit from the microplate respirometer instrument: "%"
# Time unit: "s"/"m"
# Output mass unit: "ug"
```

```{r choose_plate}
# choose a plate for downstream inspection and analysis
plate <- resp_all[[2]]
```

## 24-channel plate plots
```{r single_plate_24_channel}
channel_plot <- plate %>% 
  filter(`Time/Min.` > 15) %>% # filter out time before 15 minutes, this is usually how long the plate takes to equilibrate
  dplyr::select(`Time/Min.`, A1:D6) %>% 
  pivot_longer(cols = A1:D6,
               names_to = "Well",
               values_to = "pct_oxygen") %>% 
  ggplot(aes(x = `Time/Min.`, y = pct_oxygen)) +
  geom_point(aes(color = Well))
channel_plot
```
```{r daily_plates_24-channel}
pdf("./outputs/figures/day5_plates_24-channel.pdf")
for (i in seq_along(resp_day5)) {
plate <- resp_day5[[i]]
#
plot_title <- paste("Respirometer oxygen consumption plate", i, "24-channels")
#
channel_plot <- plate %>% 
  filter(`Time/Min.` > 15) %>% # filter out time before 15 minutes, this is usually how long the plate takes to equilibrate
  select(`Time/Min.`, A1:D6) %>% 
  pivot_longer(cols = A1:D6,
               names_to = "Well",
               values_to = "pct_oxygen") %>% 
  ggplot(aes(x = `Time/Min.`, y = pct_oxygen)) +
  geom_point(aes(color = Well)) +
  ggtitle(plot_title)
print(channel_plot)
}
dev.off()
```
```{r}
resp_all_df <- bind_rows(resp_all, .id = "id") %>% separate(id, into = c("day", "plate"), sep = "_")
```
```{r all_plates_24-channel}
raw_data_plot <- resp_all_df %>% 
  filter(`Time/Min.` > 15) %>% # filter out time before 15 minutes, this is usually how long the plate takes to equilibrate
  filter(`Time/Min.` < 60) %>% 
  dplyr::select(day, plate, `Time/Min.`, A1:D6) %>% 
  pivot_longer(cols = A1:D6,
               names_to = "Well",
               values_to = "pct_oxygen") %>% 
  ggplot(aes(x = `Time/Min.`, y = pct_oxygen)) +
  geom_point(aes(color = Well)) +
  facet_grid(plate ~ day, switch = "y")
ggsave("./outputs/figures/raw_data_plates.pdf", raw_data_plot, height = 6, width = 10)
```


## respR rate analysis
### Background respiration rate analysis
I need to calculate the background respiration rate for each plate on each day. This is also important because the blank wells appear to be gaining oxygen, so the seals may be imperfect and this requires correction.

I have the information for which wells are blanks in each run... how do I incorporate this into the rate calculation functions? Or do I do this after the rates are calculated?
```{r}
# calc_rate.bg()
```

```{r respR_single_plate}
# Use plate selected above
# Create empty data frame to hold output from rate calculation loop
rates_frame <- data.frame()
# Empty list structure to hold all rate objects
rates_list <- list()
```
```{r single_plate_respR_auto_rate}
# The below code works now! DO NOT CHANGE
# loop through rate calculation steps 
rates_frame <- data.frame("Well" = vector(), "rate" = numeric())
for (i in 3:26) {
rate <- plate %>%
  # filter(`Time/Min.` > 15 & `Time/Min.` < 45) %>% # filter out time before 10 minutes, this is usually how long the plate takes to equilibrate
  inspect(time = 2, oxygen = i) %>% 
  # Incorporate background rate detection into the analysis
  calc_rate(from = 20, to = 45, by = "time") %>%
  # auto_rate() %>% # automatically determine most linear segment
  print()  %>%   # just a quick preview
  convert_rate(o2.unit = "%",
               time.unit = "min", 
               output.unit = "ug/m", 
               volume = 0.6, S = 35, t = 26, P = 0.977)
rates_frame[i-2, 1] <- names(wells_aoi[i-2])
rates_frame[i-2, 2] <- rate$output[1]
rates_list[i-2] <- rate
}
```
```{r respR_plate_rate_functions}
# Create functions that calculate the fragment oxygen consumption rate for each well of each plate on any given day
# Different versions for calc_rate() versus auto_rate() commands
plate_resp_window <- function(plate) {
# Create empty data frame to hold output from rate calculation loop
rates_frame <- data.frame("Well" = vector(), "rate" = numeric())
# loop through rate calculation steps 
for (i in 3:26) {
rate <- plate %>%
  # filter(`Time/Min.` > 20) %>% # filter out time before 20 minutes, this is usually how long the plate takes to equilibrate
  inspect(time = 2, oxygen = i) %>%
  calc_rate(from = 20, to = 45, by = "time") %>% # calculate rate bewteen 20 and 45 minutes
  print()  %>%   # just a quick preview
   # Incorporate background rate detection into the analysis
  convert_rate(o2.unit = "%",
               time.unit = "min", 
               output.unit = "ug/m", 
               volume = 0.6, S = 35, t = 26, P = 0.977)
#
rates_frame[i-2, 1] <- names(wells_aoi[i-2])
rates_frame[i-2, 2] <- rate$output[1]
colnames(rates_frame) <- c("Well", "rate")
}
return(rates_frame)
}
# Alternate version for using the auto_rate function in respR
plate_resp_auto <- function(plate) {
# Create empty data frame to hold output from rate calculation loop
rates_frame <- data.frame("Well" = vector(), "rate" = numeric())
# loop through rate calculation steps 
for (i in 3:26) {
rate <- plate %>%
  # filter(`Time/Min.` > 20) %>% # filter out time before 20 minutes, this is usually how long the plate takes to equilibrate
  inspect(time = 2, oxygen = i) %>%
  auto_rate() %>% # automatically determine most linear segment
  print()  %>%   # just a quick preview
   # Incorporate background rate detection into the analysis
  convert_rate(o2.unit = "%",
               time.unit = "min",
               output.unit = "ug/m",
               volume = 0.6, S = 35, t = 26, P = 0.977)
#
rates_frame[i-2, 1] <- names(wells_aoi[i-2])
rates_frame[i-2, 2] <- rate$output[1]
# colnames(rates_frame) <- c("Well", "rate")
}
return(rates_frame)
}
```
### Calculate respiration rates for all plates
```{r respR_calculate_rates}
# Apply rate calculation function to all plates, returns oxygen consumption rates in units of migrograms O2 per minute.
rates_frame_auto_mapped <- resp_all %>%
  map_dfr(plate_resp_auto)
# 
rates_frame_window_mapped <- resp_all %>%
  map_dfr(plate_resp_window)
```

```{r sample_metadata_rates}
# attach metadata from keys
day_plate <- plate_key_master %>% 
  filter(Meas_Day %in% c("day2", "day3", "day4", "day5")) %>% 
  arrange(Well_ID_key)
# 
rates_frame_auto_key <- cbind(rates_frame_auto_mapped, day_plate[,1:8])
rates_frame_auto_meta <- rates_frame_auto_key %>%
  left_join(fragments, by = "Fragment_ID") %>%
  left_join(mass, by = "Fragment_ID") %>% 
  left_join(genotypes, by = "Genotype") %>% 
  mutate(rate_norm = -1*(rate / Mass))
#
rates_frame_auto_meta$Treatment <- factor(rates_frame_auto_meta$Treatment, levels = treatment_levels, ordered = TRUE)
rates_frame_auto_meta$Genotype <- factor(rates_frame_auto_meta$Genotype, levels = genotype_levels, ordered = TRUE)
rates_frame_auto_meta$Meas_Day <- factor(rates_frame_auto_meta$Meas_Day, levels = day_levels, ordered = TRUE)
```

## Inspection and exploratory analysis of respiration rates
### Blank well rates
```{r blank_rates}
rates_frame_blank <- rates_frame_auto_meta %>%
  filter(Sample_ID == "b")

rates_frame_blank %>% 
    ggplot(aes(Meas_Day, (rate)*-1)) +
  geom_violin() + geom_point() +
  coord_cartesian(ylim = c(-15, 0))

rates_frame_blank %>% 
    ggplot(aes(Plate, (rate)*-1)) +
  geom_violin() + geom_point() +
  coord_cartesian(ylim = c(-15, 0))

rates_frame_blank %>% 
    ggplot(aes(Well, (rate)*-1)) +
  geom_boxplot() + geom_point() +
  coord_cartesian(ylim = c(-15, 0))

rates_frame_blank %>% 
    ggplot(aes(Plate, (rate)*-1)) +
  geom_boxplot() + geom_point() +
  facet_grid(~ Meas_Day) + 
  coord_cartesian(ylim = c(-15, 0))
# ggsave("./outputs/physio/respR/figures/blanks_boxplots.pdf")
```
```{r daily_rates_qc}
rates_frame_samples <- rates_frame_auto_meta %>%
  filter(Sample_ID != "b")

rates_frame_samples %>% 
  ggplot(aes(Well, (rate)*-1)) +
  geom_boxplot()

rates_frame_samples %>% 
  ggplot(aes(Well, (rate)*-1)) +
  geom_boxplot() +
  facet_grid( ~Meas_Day)

rates_frame_samples %>% 
  ggplot(aes(Plate, (rate)*-1)) +
  geom_boxplot()

rates_frame_samples %>% 
  ggplot(aes(Plate, (rate)*-1)) +
  geom_boxplot() +
  facet_grid(~ Meas_Day)
ggsave("./outputs/physio/respR/figures/samples_inspect_boxplots.pdf")
```

```{r daily_raw_rates_genotype}
daily_rates_boxplot <- rates_frame_auto_meta %>% 
  filter(Meas_Day == "day5") %>% 
  ggplot(aes(Genotype, (rate)*-1, fill = Genotype)) +
  geom_boxplot(aes(color = Treatment)) +
  geom_point(shape = 21) +
  # ylim(c(0.0, 0.8)) +
  scale_fill_manual(values = genocolors) + 
  scale_color_manual(values = treatcolors[2:3]) +
  ylab("raw oxygen consumption rate (ug/min)")
daily_rates_boxplot
ggsave("./outputs/figures/daily_rates_boxplot_day5.pdf", daily_rates_boxplot, device = "pdf", height = 5, width = 10)
```

How do I normalize the rates for each individual microfragment? I can use fragment mass to start. 
How will these rates be affected by genotypic differences in morphology, density, etc.?
```{r daily_normalized_rates_genotype}
daily_norm_rates_boxplot <- rates_frame_auto_meta %>% 
  filter(Meas_Day == "day5") %>% 
  filter(Genotype != "NA") %>% 
  ggplot(aes(Genotype, rate_norm, fill = Genotype)) +
  geom_boxplot(aes(color = Treatment)) +
  geom_point(shape = 21) +
  # ylim(c(0.0, 0.8)) +
  scale_fill_manual(values = genocolors) + 
  scale_color_manual(values = treatcolors[2:3]) +
  ylab("mass-normalized oxygen consumption rate (ug/min/g)")
daily_norm_rates_boxplot
# ggsave("./outputs/figures/daily_norm_rates_boxplot_day5.pdf", daily_norm_rates_boxplot, device = "pdf", height = 5, width = 10)
```

```{r progression_normalized_rates}
rates_frame_auto_meta %>% 
  filter(Genotype != "NA") %>% 
  ggplot(aes(Genotype, rate_norm, fill = Genotype)) +
  geom_boxplot(aes(color = Treatment)) +
  geom_point(shape = 21) +
  coord_cartesian(ylim = c(0, 400)) +
  facet_grid(Meas_Day~.) +
  scale_fill_manual(values = genocolors) + 
  scale_color_manual(values = treatcolors[2:3]) +
  ylab("mass-normalized oxygen consumption rate (ug/min/g)")
```

```{r}
gulf_progression_boxplot <- rates_frame_auto_meta %>% 
  filter(!is.na(Fragment_ID)) %>%
  # filter(Gulf == "Chiriqui") %>%
  # filter(Gulf == "Panama") %>%
  ggplot(aes(Meas_Day, rate_norm)) +
  geom_boxplot(aes(fill = Genotype, color = Treatment)) +
  geom_point(aes(color = Treatment)) +
  geom_path(aes(group = Fragment_ID)) +
  facet_wrap(~Genotype + Treatment) +
  # facet_grid(Treatment~.)+
  scale_fill_manual(values = genocolors) + 
  scale_color_manual(values = treatcolors[2:3])
gulf_progression_boxplot
# ggsave("./outputs/figures/gulf_respiration_boxplot.pdf", height = 8, width = 10, device = "pdf")
```

```{r}
rates_geno <- rates_frame_auto_meta %>%
  group_by(Treatment, Genotype, Meas_Day) %>% 
  summarise(geno_rate_norm = mean(rate_norm)) %>% 
  filter(!is.na(Genotype))

geno_rates_plot <- rates_frame_auto_meta %>% 
  filter(Genotype != "NA") %>% 
  ggplot(aes(Meas_Day, rate_norm)) +
  geom_boxplot(aes(fill = Treatment)) +
  geom_path(data = rates_geno, aes(Meas_Day, geno_rate_norm, group = Genotype, color = Genotype)) +
  # geom_smooth(aes(Meas_Day, rate_norm, group = Genotype, color = Genotype)) +
  geom_point(data = rates_geno, aes(Meas_Day, geno_rate_norm, color = Genotype), size = 5) +
  coord_cartesian(ylim = c(0, 400)) +
  facet_grid(~Treatment) +
  scale_color_manual(values = genocolors) + 
  scale_fill_manual(values = treatcolors[2:3]) +
  xlab("Measurement Day") + 
  ylab("mass-normalized oxygen consumption rate (ug/min/g)") + 
  ggtitle("Mass-normalized oxygen consumption rates during experiment")
geno_rates_plot
ggsave("./outputs/physio/respR/figures/geno_rates_boxplot.pdf", height = 5, width = 10, device = "pdf")
```

## Statistical analysis
```{r}
rates_frame_auto_meta1 <- rates_frame_auto_meta %>%
  filter(Sample_ID != "b") %>% 
  mutate(lrate = log10(rate_norm),
         lnrate = log(rate_norm, base = exp(1)),
         invrate = 1/rate_norm)

shapiro.test(rates_frame_auto_meta1$rate)
ggdensity(rates_frame_auto_meta1$rate)
ggqqplot(rates_frame_auto_meta1$rate)

shapiro.test(rates_frame_auto_meta1$rate_norm)
ggdensity(rates_frame_auto_meta1$rate_norm)
ggqqplot(rates_frame_auto_meta1$rate_norm)

shapiro.test(rates_frame_auto_meta1$lnrate)
ggdensity(rates_frame_auto_meta1$lnrate) + stat_overlay_normal_density(color = "red", linetype = "dashed")
ggqqplot(rates_frame_auto_meta1$lrate)

shapiro.test(rates_frame_auto_meta1$invrate)
ggdensity(rates_frame_auto_meta1$invrate) + stat_overlay_normal_density(color = "red", linetype = "dashed")
ggqqplot(rates_frame_auto_meta1$invrate)
```
```{r}
rates_frame_auto_meta1 %>% 
  group_by(Genotype) %>% 
  shapiro_test(rate_norm, lnrate) %>% filter(p < 0.05) %>% View()
# %>% group_by(variable) %>% count()
```
```{r}
shapiro.test(mass$Mass)
ggdensity(mass$Mass)
```
```{r}
kruskal.test(rates_frame_auto_meta1)
```
```{r}
fit <- lm(rate_norm ~ Treatment + Genotype + Meas_Day + Treatment*Genotype, data = rates_frame_auto_meta1)
summary(fit)
# 
# Other useful functions
coefficients(fit) # model coefficients
confint(fit, level=0.95) # CIs for model parameters
fitted(fit) # predicted values
residuals(fit) # residuals
anova(fit) # anova table
vcov(fit) # covariance matrix for model parameters
influence(fit) # regression diagnostics
# diagnostic plots
layout(matrix(c(1,2,3,4),2,2)) # optional 4 graphs/page
plot(fit)
# 
shapiro.test(residuals(fit))
```
```{r}
# Stepwise Regression
library(MASS)
fit <- lm(rate_norm ~ Treatment + Genotype + Meas_Day, data = rates_frame_auto_meta1)
step <- stepAIC(fit, direction="both")
step$anova # display results
```
```{r}
# Global test of model assumptions
gvmodel <- gvlma(fit)
summary(gvmodel)
```
```{r}
model_matrix(rates_frame_auto_meta1, rate_norm ~ Treatment + Gulf + Genotype + Meas_Day) %>% View()
```

### Mixed-effect linear models
Apply linear mixed effects models to the respiration data akin to the IPAM data...
```{r}
rates_frame_auto_meta
rates_frame_auto_meta$Day<-as.integer(rates_frame_auto_meta$Meas_Day)
```
```{r}
rates_frame_samples <- rates_frame_auto_meta %>% filter(Sample_ID != "b")
```

```{r}
Resp_Genotype <- ggplot(rates_frame_samples, aes (Meas_Day, rate_norm, color = Genotype)) +
  geom_point(alpha=0.3) +
  stat_summary(fun.data = "mean_cl_boot", geom = "errorbar", width = 0.2 ) +
  stat_summary(fun.y=mean, geom = "line") +
  scale_color_manual(values = genocolors) +
  theme_classic()

Resp_Genotype + facet_grid(ITS2~Treatment)
Resp_Genotype + facet_grid(mtORF~Treatment)
Resp_Genotype + facet_grid(Location~Treatment)
```
```{r}
# Treatment 
Resp_Treatment <- ggplot(rates_frame_samples, aes (Meas_Day, rate_norm, colour=Treatment)) +
  stat_summary(fun.data = "mean_cl_boot",geom = "errorbar", width = 0.2)+
  #stat_summary(fun.y=mean, geom="point", size =3, alpha=0.5) +
  stat_summary(fun.y=mean, geom="line") + theme_classic()+
  theme(plot.background=element_blank(), 
            panel.border = element_blank(),
            axis.line = element_line(colour = "black"),
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            legend.position="bottom",
            strip.background = element_rect(fill="white")) +
      scale_color_manual(values = treatcolors[2:3]) #+
      # scale_y_continuous(limits = c(0.1, 0.7),
      #                    breaks = seq(0.2, 0.6,0.2),  
      #                    expand = c(0, 0),
      #                    name=("YII (Fv/Fm)")) +
      # scale_x_continuous(name="Days in the experiment",
      #                    limits = c(0,6),
      #                    breaks = seq(0, 6, 1),  
      #                    expand = c(0, 0))
Resp_Treatment
Resp_Treatment + facet_wrap(~mtORF)
Resp_Treatment + facet_wrap(~ITS2)
```
```{r}
# Libraries 
  library("lme4")
  library("multcomp")
  library("multcompView")
  library("emmeans")
  library("effects")
  library("lmerTest")
```

If you use day as a factor:

* mtORF and ITS2 considered random effects
* Day (as factor, not numeric) and Treatment considered fixed effect (what you are really testing)


```{r}
# ipam_fragments$DayF<-as.factor(ipam_fragments$Day)

LM_1 <- lmer(rate_norm ~ Treatment * Meas_Day + 
                             (1|mtORF) + (1|ITS2), REML=TRUE, data= rates_frame_samples)
anova(LM_1)
ranova(LM_1)
step(LM_1)
# * fv/Fm is significantly different (not a surprise given day 2 data)
# * No treatment effect
# * Both, ORF and ITS2 are significant
```
```{r effect_plot}
pdf("./outputs/physio/respR/figures/respR_treatment_effect.pdf", width = 8, height = 6)
plot(Effect(c("Treatment","Meas_Day"), LM_1), x.var="Meas_Day", multiline = T, ci.style="bars")
# Clear that there is no treatment effect after controlling for other random factors
dev.off()
```
```{r}
Effect(c("Treatment","Meas_Day"), LM_1)
```

# To look at the ITS2, ORF, etc

This is kind of meaningless until we close the gaps in ITS2 and ORF, other option is discard genotypes with "unknown" and run with the genotyped corals.  


```{r make ITS2 the question}
rates_frame_samples_known <-  rates_frame_samples %>% filter(Genotype != "PAN-5" | Genotype != "PAN-10" | Genotype != "URA-51")

LM_2 <- lmer(rate_norm ~ ITS2 + mtORF + 
                             (1|Meas_Day), REML=TRUE,  data = rates_frame_samples_known)
anova(LM_2)
ranova(LM_2)
step(LM_2)
#  Pair-wise comparisons
cld(emmeans(LM_2, "mtORF"))
RespEmm <- cld(emmeans(LM_2, specs = c("ITS2")))
#write.csv(YIIAcerEmm, "YIIAcerEmm.csv")
RespEmm
```
